---
title: "Exercícios - Visualização de Dados (Capítulo 1)"
author: "Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: cerulean
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Exercícios - Pág. 6

## Run `ggplot(data = mpg)`. What do you see?

```{r}
library(tidyverse)
ggplot(data = mpg)
```

Vemos uma tela cinza, sem eixos e sem gráfico.

## How many rows are in `mtcars`? How many columns?

```{r}
dim(mtcars)
```

A base de dados `mtcars` tem 32 linhas e 11 colunas.

## What does the `drv` variable describe? Read the help for `?mpg` to find out

```{r, include=FALSE}
# ?mpg
```

-   `drv`: The type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

-   Trem de força ou transmissão do carro (tração): dianteira, traseira e 4 x 4.

-   Pacote de tradução das bases de dados do R.

```{r}
# install.packages("remotes")
# remotes::install_github("cienciadedatos/dados")
```

```{r}
library(dados)
glimpse(milhas)
```

## Make a scatterplot of `hwy` versus `cyl`.

```{r}
mpg %>% 
  ggplot(aes(x = cyl, y = hwy)) +
  geom_point()
```

```{r}
class(mpg$cyl)
```

## What happens if you make a scatterplot of `class` versus `drv`? Why is the plot not useful?

```{r}
mpg %>% 
  ggplot(aes(x = drv, y = class)) +
  geom_point()
```

Porque temos duas variáveis categóricas. Gráficos de dispersão são mais úteis para verificar relações entre variáveis numéricas.

# Exercícios - Pág. 12

## What's gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

## Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?

## Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical versus continuous variables?

## What happens if you map the same variable to multiple aesthetics?

## What does the stroke aesthetic do? What shapes does it work with? (Hint: use `?geom_point`.)

```{r}
library(ggplot2)
?geom_point
vignette("ggplot2-specs")
```

-   `Stroke` é a largura da marca ao redor do ponto.

```{r}
sizes <- expand.grid(size = (0:3) * 2, stroke = (0:3) * 2)
```

-   Aqui, o `stroke` é mapeado pela variável `stroke`, ou seja, quanto maior o valor, maior a marca em torno do ponto.

```{r}
ggplot(sizes, aes(size, stroke, size = size, stroke = stroke)) + 
  geom_abline(slope = -1, intercept = 6, colour = "white", size = 6) + 
  geom_point(shape = 21, fill = "red") +
  scale_size_identity()
```

-   Alterando os valores da estética `stroke`:

```{r}
ggplot(sizes, aes(size, stroke, size = size, stroke = 2)) + 
  geom_abline(slope = -1, intercept = 6, colour = "white", size = 6) + 
  geom_point(shape = 21, fill = "red") +
  scale_size_identity()
```

-   Note that shapes 21-24 have both stroke `colour` and a `fill`.

-   A estética `stroke` só funciona com as formas de 21 a 24.

-   Mais um exemplo:

```{r}
library(tidyverse)
glimpse(mpg)
```

```{r}
iris %>% 
  ggplot(aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point(size = 2, 
             stroke = 1, 
             shape = 21, 
             fill = "yellow",
             color = "red")
```

## What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`?
